
# Methods: Model testing and calibration

## Individual level variable inclusion

From the list of variables that could have an influence in Plasma Se, we decided to include the following:

- Age
- Residency
- Wealth quintile

Based on the previous analysis performed in the same dataset by Phiri et al (2019). In addition, based on the literature (Combs, 2015, Combs XX, ...), we explore the inclusion of the following parameter:

- Inflammation 
- Smoking
- BMI
- Malaria

Of those four, we only included inflammation, smoking was excluded due to the low number of women reporting smoking in our sample, and the low prevalence of smoking behaviour among women in Malawi (see variable exploration section for more information), BMI was excluded as most of the women were within normal weight range, and there was no strong correlation between BMI and Plasma Se, and also it has only be reported to be correlated to SEPP1 in literature, and studies done in high income countries, where Se intake and plasma level as well as weight and BMI were higher than in our population. Finally, malaria was excluded because it has low correlation with the plasma Se (no significant difference between the two groups) and similar prevalence was reported across regions, and we because its impact it may be well explained by inflammation. 


## Geostatistics & point data

INLA + SPDE 

### INLA resources

Tutorial & good visualisation
https://ourcodingclub.github.io/tutorials/inla/

## Parameter inclusion


The covariates and the rationale for their selection are detailed in section XX. In summary, all the covariates included as potentially explanatory variables for individual variability in plasma Se. The first model to be fitted included all of the covariates plus an spatial random effect modelled using SPDE (@ref-).


```{r}

# Formula for the model
form1 <- log(y) ~ -1 + Intercept +  log(Se_mean) +
  wealth_quintile + BMI +  urbanity + 
  AGE_IN_YEARS +
  log(crp) + log(agp) + f(spatial.field, model = spde) 


```


Then, we tested the same model but excluding the urbanity, and only including rural household. 

```{r}

# Formula for the model
form2 <- log(y) ~ -1 + Intercept +  log(Se_mean) +
  wealth_quintile + BMI +  urbanity + 
  AGE_IN_YEARS +
  log(crp) + log(agp) + f(spatial.field, model = spde) +
  f(ID, model = 'iid')


```

Then, we added the cluster as random effect

## Model 3

We added the variable distance to water body (`dist_to_wb`) to the model. This variable was not normally distrubuted following the skewerness and ocktile skerness ranges. Hence, we needed to log tranform and hence we decided to change the 0 values to 200. On the basis that this distance was smaller than the min. distance and smaller than the displacement (2.5km smalles for urban), as it is impossible that the distance to a water body is 0m. 

Then we tested againt the 10aggregation unit. Again, the main conclusion remained the same, hence only age and maize Se concentation were significantly associated with plasma Se concentration. However, adding distance to the water body improved the overall performance of the models. 

## Testing Model 4

We removed the spatial random effect, using only the iid model using the cluster as random effect. 

Overall model performance was good (-100 to -102), but again no major differences between the aggregation levels. 

We saw bigger effect on the maize Se (although check whether is only the scale), the distance to lakes (same as before), and the intercept. 

Question: Is the intercept indicating that the spatial is performing better (high non-accounted variability)??

Also, although non-significant we could see a pattern for the wealth Q, with the low-est Q negative influencing plasma Se and highest positively.

Question: Check whether urbanity and cluster may have interaction effect. 


# Testing 5

When we used only the predicted values with buffers from 10-60km and cluster and district. The best performing model was the one using the cluster. Interenstingly, we can see a decrease in the model fit as the buffer increases. Also, the distance to the lake, was only negatively correlated with some of the buffers aggregation (e.g, 2-4, and then 7and 8). 

These results were obtained when establishing prior. 

**TODO**: Check if holds w/o the priors. 


# Results

The four models for the 10 aggregation units performed similarly, with the main conclusion being the same. Maize Se and age are important covariates irrespectively of the aggregation level. 




# Model assumptions

### Model testing

#### INLA SPDE

First we select the data, we are testing 10 levels of aggregation as explained in sec. XX, which are the following:

`r print(file)`

We are then applying the same model to the 10 different level of aggregation, so we create a loop that perform the individual steps for all 10 datasets and stores the model output. 


## The mesh

We are have tested few mesh sizes and can be seeing in the `mesh-test.R`. This steps is crucial as stated in several studies (REF, REF, REF).

To create the mesh we used the function `inla.mesh.2d()` and it can be seen here.

```{r}

# Locations
coord <- cbind(plasma_se$Longitude, plasma_se$Latitude)

#Creating the mesh
mesh <- inla.mesh.2d(loc = coord, 
                     max.edge = c(.5, 3), 
                     cutoff = c(0.1))

plot(mesh)
points(coord, col = "red", pch = 2)

```




# Model selection


# Model validation & checks


## Sensitivity

The model is robust as the conclusion remained the same through out the three model tested, 

### Household displacement

In order to see how the displacement may potentially affect the model results. We have displaced all/ X% of the sample within range. Hence, the point locations were displaced by 2km, and the analysis were re-run. 







