---
title: "Reproducible data documentation for the study on geospatial relationship between plasma Se concetration in WRA & maize grain Se concentration in Malawi"
format: 
  html:
    number-sections: true
    number-offset: 1
    code-fold: show 
    embed-resources: true
editor: source
---

```{r set-up, message=FALSE, warning=FALSE}

# Loading packages
library(dplyr) # data wrangling 
library(sf) # spatial data manipulation

# Data
# Biomarkers data DHS (00_cleaning-dhs, line 25)
Malawi_WRA <- haven::read_dta(here::here("data", "MW_WRA.dta")) 

# Survey analysis: Cleaned dataset weight (00_cleaning-dhs, line 574)
EligibleDHS  <- readRDS(file=here::here("data", "inter-output","dhs_se_gps.rds"))

# Loading maize data (Kumssa - GeoNut, 00_cleaning-maize, line 16)
data.df  <- read.csv(here::here("data", "maize", "MWI_CropSoilChemData_CSV", # Grain & soil chem data
 "MWI_CropSoilData_Raw.csv"))

# Loading maize data (complementary) (Chilimba, 00_cleaning-maize, line 16)
maize  <- readxl::read_excel(here::here("data", "maize",
 "AllanChilimbaFieldData_forLucia_20230615.xlsx"), sheet = 1)

# Admin Boundaries for Malawi 

# EAs
ea_bnd  <- st_read(here::here("data", "mwi-boundaries", "EN_NSO" , "eas_bnd.shp"), quiet = TRUE)

# TAs
ta_bnd  <- st_read(here::here("data", "mwi-boundaries",
 "mwi_adm_nso_hotosm_20230329_shp","mwi_admbnda_adm3_nso_hotosm_20230329.shp"), 
                   quiet = TRUE)


# Loading stat functions
source(here::here("functions", "CEPHaStat_3.R"))


```



## Introduction {.unnumbered}

### Objectives {.unnumbered}

The objectives of these documentation and the exploratory analysis are:

1)  to document the cleaning steps and decision made that may influence the analysis and subsequent results.

2)  to evaluate the different variables that will be used in the analysis

3) to document and calculate the mean/median maize Se concentration at different level of spatial aggregation (i.e., from 5km radius up to regional averages).

## Dataset & Variables 

In this section we explained the datasets and variables that were evaluated for use in the study: "{Insert title of the study}". We provided information about the data provenance and the reasons for inclusion/exclusion in the analysis. 


### Geo-referenced crop (maize) composition (Se concentration) data in Malawi

#### Main dataset: GeoNutrition

This dataset was obtained from this [open repository](https://doi.org/10.6084/m9.figshare.15911973), and it is described by Kumssa et al, (2022) & Gashu et al, (2021). We decided to use this dataset instead of the one reported in the repository accompanying Gashu and colleagues study (2021) (see excluded data sources below) because it provided the raw dataset (i.e., including below detection limit values), as well as more metadata.

The data repository contains a number of data files for [Malawi](https://figshare.com/articles/dataset/Cereal_grain_mineral_micronutrient_and_soil_chemistry_data_from_GeoNutrition_surveys_in_Ethiopia_and_Malawi/15911973?file=35796662):

-  **Raw crop mineral dataset** (MWI_CropSoilData_Raw), which provides the raw values, as extracted from the analytical device (i.e, data from the ICP without any further processing). 
 
-  **Processed crop mineral dataset** (MWI_CropSoilData_NA), which has turned values which were not analysed (NM), below the limit of detection (BLOD), or negative values into `NA`.  

-  **Limit of detection dataset** (MWI_Crop_LOD_ByICPRun), it provides information about the limit of detection per each round of ICP.


This is problematic for our analysis as selenium (Se) in Malawi was highly affected due to the very low values of Se concentration in maize in that country. Therefore, some considerations were taken and are explained in the data exploration (@sec-explo) as well as in the modelling section (Section 3.x). See also notes in Kumssa et al, (2022). 

In addition, we found that in Southern Malawi were other crops (e.g., millet, sorghum, etc.) were collected instead of maize. It could be because the harvest season is slightly earlier in that area and hence the fields were already plotted and other crops planted (@ref).

File name: `data/maize/MWI_CropSoilChemData_CSV/MWI_CropSoilData_Raw.csv`

File name: `data/maize/MWI_CropSoilChemData_CSV/MWI_Crop_LOD_ByICPRun.csv`

#### Complementary dataset: Chilimba 

Malawi maize grain and soil Se concentration dataset (2009-2010): It is publicly available [here](), but we obtained a more extensive dataset from the co-authors of the manuscript (E.L.A.) including the values of pH (in water) of the soil where samples were collected. 

This dataset was included to provide information on maize grain Se concentration in Southern Malawi. 

File name: `data/maize/AllanChilimbaFieldData_forLucia_20230615.xlsx`


#### Excluded sources of data


  a. Malawi maize grain and soil Se concentration dataset (2018). It is publicly available [here](https://github.com/rmlark/GeoNutrition/tree/main/Soil_Crop_comparisons/Malawi)

  File name: `GeoNutrition/Soil_Crop_comparisons/Malawi/Malawi_grain_soil.xlsx` 
(data/maize/Malawi_Grain.csv)


  b. Malawi crops, including maize, mineral concentration dataset (2012-2013) and soil pH. It is publicly available [here](https://www.sciencedirect.com/science/article/pii/S0048969714014764). (Joy et al., 2014)(@ref). 

  File source: [Supplemantary table](https://ars-els-cdn-com.ez.lshtm.ac.uk/content/image/1-s2.0-S0048969714014764-mmc1.xlsx)
  
  File re-name: `data/maize/2012_Joy-mwi-samples.xlsx`

**Tabs:1-10**: Explored in this analysis:

-   STable6: crop sample info (maize Se) and coordinates
-   STable5: Soil samples (soil pH)
-   STable10: Information on each crop and its corresponding soil sample (Sample_number, Soil_sample).

**Cleaning script:** excluded/00_cleaning_excluded.R

For more information regarding the exclusion go to Section 2 (*Cleaning and Exploratory analysis*).

#### Enviromental covariates (Mean Annual Temperature)

This raster dataset (mwi_CHELSA_bio10_1.tif) contains information on the Mean Annual Temperature (MAT) for Malawi from 1979-2013. 

The dataset is a cropped (clipped) area from the global (CHELSA_bio10_1.tif) which is a worldwide high resolution raster layer. The geographic coordinate system are referenced to the WGS 84 horizontal datum, with the horizontal coordinates expressed in decimal degrees. 
The dataset is in GEOtiff format and was extracted from the CHELSA dataset [here](https://datadryad.org/stash/dataset/doi:10.5061/dryad.kd1d4). This variable was used as environmental covariates for the geospatial Se maize model (`BIO1`). 

The clipping was done on QGIS version (3.28.3) Firenze using the ea_bnd layer as reference extend for clipping. 

Sequence: Raster>Extraction>Clip Raster by Extend...>Clipping extent>Calculate from a layer> ea_bnd. 

### Geo-referenced biomarkers data (Plasma Se conc. in WRA) in Malawi 

<br>

Malawi DHS - MNS 2015-2016. Access need to be requested to DHS data [here](https://dhsprogram.com/methodology/survey/survey-display-483.cfm).


#### Micronutrient survey data (MW_WRA)


This dataset provided information on biomarkers and other biological data for women of reproductive age in Malawi. The recode and all information about the variable description can be found [here](~\MW_2015-16_DHS_09262022_949_164775.zip\MWOB7ASV). Here's the selection of variables that were considered for cleaning and exploration.

-   **mregion** (region) [1-3 = N->S]
-   **mtype** (urban/rural) [U-R = 1-2] 
-   **m04**  (gender)
-   **mcluster** (cluster number - PSU)
-   **mnumber** (husehold number - SSU)
-   **mweight** (Micro nutrient survey household weight (6 decimals))
-   **m01** (line number - individual ID within HH)
-   **m07** (age) (age in years)
-   **m435** (weight)
-   **m436** (height)
-   **sel** (Plasma Se (ng/ml))
-   **m415** (Took any kind of vitamin or mineral tablet/syrup/powder)
-   **m431** (Malaria test result)[^1]
-   **agp_crp_c1** (ANY_INFLAMMATION, Any inflammation (AGP >1 g/L or CRP >5 mg/L)
-   **agp** (AGP 'Alpha-1-Acid Glycoprotein (g/L), (inflammation biomarker))
-   **crp** (CRP C-Reactive Protein (mg/L), (inflammation biomarker))

[^1] Malaria test recode: Final result of malaria from blood smear test  

-  2 = Negative (in recode manual = 0)
-  1 = Positive
-  6 = Test undetermined (4)
-  7 = Sample not found in lab database (6)
-  9 = Missing (m)
-  (na) = Not applicable


#### Questionnaire data (MWIR7) 

This dataset provided information on socioeconomic factors for the whole datasets. We downloaded both the .dat and .dta file because the file with .dat extension was not opening in R. It was not corrupted as we could open the .MAP with Notepad+, and we could see all recode documentation. 

These variables provide information at individual level and, below are the variables of interest:

-   **v101/V024** (De facto region of residence)
-   **sdist** (district). 
-   **v022** (survey strata)[^2].
-   **v005** (Woman's sample weights, divide it by 1000000 before applying the weighting factor). Not to be used for MNS, as MNS is a subset and has its own survey weights. 
-   **v106** (Highest educational level)
-   **v155** (Literacy)
-   **v190** (Wealth index)
-   **v190a** (Wealth index for urban/rural) 
-   **v002** (Household number)
-   **v001** (Cluster number)
-   **v463a** (is_smoker, Only cover cigarettes (other smoking variables))
-   **m420** (had_malaria)
-   **v150** (Relationship to household head)

[^2]Note that v022 & v023 are strata Urban/rural district as the information collected here was for the whole DHS survey. This should not be used for MNS because it is a subset of the DHS survey with the representatives (and strata) in at region+rural/urban. We could create our own strata based on the survey design (1/6 (3regions*2urban/rural)).

#### Location (GPS) data (MWGE7AFL)

This dataset provides information on GPS coordinates (displaced) for each cluster (PSU = EA). Hence, all the households in the same EA/cluster have the same (GPS) location.

The variable used for merging is DHSCLUST (to be merged with V001).  

Note that GPS displacement is true within district. Also, according to the documentation the cluster "centre" recorded is the "center of the populated place in the cluster" (MWGE7AFL/GPS_Displacement_README.txt, DHS, 2014).


### Boundaries of Malawi {#sec-bound}

Malawi administrative boundaries are, from the largest to smallest area: Region> District> Traditional Authorities (TA)> Enumeration Areas (EA). We are mainly interested in Districts and EAs for our modelling study. Other boundaries in Malawi that were explored included: agro-ecological zones, agricultural land, body water, national parks, etc. 

We excluded Agro-Ecological Zones (AEZs) boundaries from our study because there are only 4 AEZs in Malawi, hence areas are too large to get any meaningful spatial pattern, and/or level of aggregation. 

Note: The AEZs shape files can be found in the Scoping review (R project, data folder), and further information can be found in [here](https://climateknowledgeportal.worldbank.org/sites/default/files/2019-06/CSA%20_Profile_Malawi.pdf).

Information of the boundaries were found from difference sources and data owners. Here's a list of the boundaries we explored and the reasons for inclusion/ exclusion. 

#### Main administrative boudaries file

3. [Population over Enumeration Areas NSO, (eas_bnd)](https://www.masdap.mw/catalogue/#/dataset/126): this layer shows the distribution of the population over the country of Malawi on the basis of enumeration area boundary. Data producer is the National Statistical Office.

File: `mwi-boundaries/EN_NSO/eas_bnd.shp`


#### Excluded boudaries files


1. [GADM maps and data (gadm40_MWI_shp)](https://gadm.org/download_country.html): The file for the EAs has some issues that led to some data points falling into the lake. Hence, we found a different file with the Malawi boundaries. This is because some of the EAs near the lake has some "ownership" over certain areas of the lake, which were included in the boundaries of those EAs in the shapefile (i.e., as a continuum from land to lake).   

2. [Malawi NSO from HDX (Humanitarian datasets) (mwi_adm_nso_hotosm_20230329)](https://data.humdata.org/dataset/cod-ab-mwi). This source only have up to Admin2 (District) level. 

4. Echo2 prioritization, these files covered up to EAs. (depricated - Updated version: Population over Enumeration Areas NSO).

5. GeoBoundaries: It also has 3 level of boudaries: 1-region, 2-district, 3-TA. These files has no EA boundaries (deprecated).

Other excluded boundaries: Agro-Ecological Zones, Traditional Authorities (more information on data exploration: @sec-explo-bound).


### Geo-referenced apparent food consumption in Malawi (Se intake) - REMOVE?

This dataset was only used as support information. Fourth Integrated Household Survey 2016-2017. https://doi.org/10.48529/g2p9-9r19. It can be accessed through the World Bank microdata website. 

Information about each module and the information contained can be found [here](https://microdata.worldbank.org/index.php/catalog/2936/pdf-documentation)

**File of interest & variables**:

a. Food consumption

b. Geographic data (HouseholdGeovariablesIHS4.csv). This dataset has information on the displaced coordinates of each household, distance to some features (road, population centre, agricultural market, etc), and other enviromental variables (elevation, rainfall, etc).

Variables of interest:

-   case_id: Unique Household Identifier

-   HHID: household id (Survey Solutions Unique HH Identifier)

-   lat_modified: GPS Latitude Modified

-   lon_modified: GPS Longitude Modified

-   dist_road: HH Distance in (KMs) to Nearest Road

-   dist_agmrkt: HH Distance in (KMs) to Nearest Agricultural Market

-   ssa_aez09: Agro-ecological Zones

-   srtm_1k: Elevation (m)

c. Sales vs stored crops (rainy season) (AG_MOD_I)
d. Sales vs stored crops (dry (dimba) season) (AG_MOD_O)
d. Sales vs stored trees (12months) (AG_MOD_Q)



## Cleaning & Exploratory Analysis {#sec-explo}

### Maize grain Se concentration in Malawi

#### GeoNutrition (Kumssa et al., 2022)

We extracted multiple variables of interest (including the LOD), which served to: 

a)  check the `NA` values (n=427) for maize Se concentration in found in the dataset published in Gashu et al. study (2021).

b)  to add those values as possible solution for accounting for the high number of missing maize Se conc. values.

Due to the high number of BLOD we have generated four Se variables to test sensitivity of the model.

 - **Se_raw**: This variable contains the values as reported from the chemical analysis including the values BLOD. Cleaning steps: 
 
i)  No measured: Six Se values were coded as not measured ("NM") , which were transformed into `NA`. 

```{r}

subset(data.df, ID %in% c("MWI0574", "MWI1007","MWI1138","MWI1168","MWI1171","MWI1686"), 
       select = c(ID, Crop, Se_grain)) %>% knitr::kable()

```

ii)  Negative (or zero) values: 21 values were negative and converted to the smallest limit of detention value (0.002..). We tested to convert it into: zero, which is a problem for a number of reasons, including that zero cannot be log transformed (-Inf.), and to the minimum value in the dataset (0.0009) which again generate problems with outliers, and skewerness.  

 - **Se_std**: This variable contains Se values as above, however BLOD values were substituted with their respective LOD value (as per each round of test, see `Crop_ICP_Run_Se` & `Se_LOD` variable).
  
- **Se_grain**: This contains the Se values reported above the LOD, where values BLOD were excluded (i.e., converted into `NA`), as was previously done by Gashu et al., 2021 and reported in the "MWI_CropSoilData_NA.csv" dataset in Kumssa et al, 2022. 


Additionally, the MAT values were added to this dataset for each sample location (i.e., GPS coord.), as this study only reported analytical values carried as part of the GeoNutrition project. These values were completed when extracting the MAT for the Chilimba et al. (2011) sample locations. 


##### Spatial preliminary analysis


The missing (i.e., BLOD values) are spatially clustered, which probably will influence (unbalancing) our results (Kumssa et al, 2022). We hypothesised that this is due to high concentration of very low maize grain Se conc. values likely driven by soil and environmental factors in those affected regions(@ref). 

Although, the samples collected provided a good coverage of Malawi, there were 2200 EAs without a maize Se conc. samples. In addition, low number of maize samples were found in the southern region of Malawi (Shire Valley), leading to a more scarce and, probably higher uncertainty in that region.  Moreorver, consumption of maize in that region is comparable with the other regions of Malawi, and probably the main reason for the lower maize values is that the harvest season is slightly earlier than in other regions of Malawi (probably due to difference in the climatic patterns), and hence maize was not found when the field team were there to collect the data. 

In order to minimise the impact on our model, and after checking that maize consumption is the same as other regions in Malawi (even slightly higher), we decided to explore other datasets which led to the inclusion of maize samples from Chilimba et al. study (2011) (see data section & next section).


#### Mineral concentration in maize (Chilimba et al., 2011)


This dataset provided Se concentration in maize grain and pH in soil (measured in water) in the same locations, however it did not contain information on the MAT (`BIO1`) for each location, which are used for the maize aggregation model (see _02modelling). 

We extracted the variable of interest, removed `NA` and checked the coordinate projections to make sure both datasets (GeoNutrition & this dataset) were compatible (both datasets provided longitude and latitude data in decimal degrees, WGS84 datum, coordinate reference system EPSG:4326). 

```{r, cleaning-steps}

#Renaming variable & adding info on data source
names(maize)[c(23, 28)]  <- c( "Se_mg", "pH_w")
maize$survey  <- "Chilimba"

# Getting variables of interest 1:12 (sample descript), grain_se and pH
maize  <-  maize[,c(1:11, 23, 28, 56)]

# Checking values 
sum(is.na(maize$Se_mg)) # Only 2

```


When we checked the distribution of the sample (@fig-2a, @fig-2b), we can see that the log-distribution shape is a bit odd. This may be related to the majority of the samples were collected from high maize Se conc. areas (i.e., Shirey valley), which, paired with probably the omission of BLOD values, led to a large number of high concentration values. 

```{r label = "fig-2a", fig.cap = "Distrubution of maize Se concentration from the study of Chilimba et al. (2011)."}

summaplot(maize$Se_mg)

```


```{r label = "fig-2b", fig.cap = "Distrubution of log transformed maize Se concentration from the study of Chilimba et al. (2011)."}

summaplot(log(maize$Se_mg))

```


#### Excluded datasets

##### GeoNutrition (Gashu et al., 2021)


This dataset was excluded because high percentage of values were transformed to `NA` some of which were BLOD values for maize Se concentration. However the cleaning scripts are provided in this repository. 

We extracted the variables of interest from the dataset: Se conc. (only in maize), pH, MAT, and the coordinates. Then we are doing some data operations, for instance, removing NAs (values >999) in Se conc. and checking missing values for the covariates. 

Two pH values were missing, hence we may exclude those values for the modelling, also there were three locations with missing MAT. This could be have been completed when extracting the MAT for the Chilimba et al. (2011) sample locations. 


##### Mineral concentration in crops (Joy et al., 2014)

This study was excluded because it did not provided substantial information (i.e, only few new samples), while it would potentially add more uncertainity to the final results. The script for cleaning the dataset is provided in this repository.

This study collected data for various crops, hence we needed to clean the dataset to obtain only the information for maize Grain. We also excluded maize flour to keep it consistent with our dataset. In addition, soil pH values were extracted to be used as model covariates. 

One limitation of the dataset is that the GPS location were only provided for crops and not for soils, hence we were trying to use some information (STable10) to combine the sample number with the soil samples. These led to the correct allocation of a pH value for 57 of the 155 samples. 

According to the documentation, soil samples were not taken when the samples were collected at the market. We are then assuming that samples without pH value were collected at the market and hence should be excluded from our analysis. Then, only one or two samples were collected in the Shire Valley, thus this sample dataset will not be contributing substantially to reduce the uncertainty, and it may add other sources of variability (i.e., temporal, analytical, etc). Hence, it was excluded from the final dataset.

#### Mean Annual Temperature

Mean Annual Temperature (MAT) was used as environmental covariate for the aggregation of maize Se model. The data was extracted from the CHELSA dataset which was downloaded from [here](https://datadryad.org/stash/dataset/doi:10.5061/dryad.kd1d4) on 2023/07/17 and cropped for Malawi for easy usage. The geographic coordinate system are referenced to the WGS 84 horizontal datum, with the horizontal coordinates expressed in decimal degrees.  

#### The combined dataset: Crop Se dataset

When we combined both GeoNutrition (Kumassa et al, 2022) and Chilimba datasets, as expected, the maize Se concentration is heavily skewed and it would need log transformation, while pH has a normal distribution. 

For more information on that see the exploratory analysis and the variable selection for the model reported in Gashu et al., (2019) [@ref] and the analysis done by Lark et al. (2019), here are the [scripts](https://github.com/rmlark/GeoNutrition/blob/main/Soil_Crop_comparisons/Malawi/Malawi_model.R). 

#### Limitations of the maize data

- **Low data availability in Southern Malawi region**. For instance,in our main dataset, the GeoNutrition, mainly other crops (e.g., sorghum, millet, etc.) were collected (~11%).
  
- **High proportion of below limit of detection values** (~25%). Risk of bias:
    - Potential over-estimation (or under-) of maize Se concentration
    - Not at-random missing values --> Spatial pattern


## Se biomarker dataset (DHS-MNS)

### Micronutrient survey data (MW_WRA)

```{r collapse=TRUE, cleaning-mw_wra}

# Renaming variables 
Malawi_WRA <- Malawi_WRA %>% 
  dplyr::rename(
  ferritin='fer',
  region='mregion',
  sex = "m04", 
  zinc='zn_gdl',
  dretinol='incap_dr',
  retinol='incap_retinol',
  urbanity='mtype',
  MRDR='mrdr_ratio',
  survey_cluster1='mcluster',
  household_id1='mnumber',
  survey_weight="mweight",
  LINENUMBER='m01',
  AGE_IN_YEARS='m07',
  SALT_PRESENT='m12',
  SALT_LABEL_IODIZED='m104',
  OIL_PRESENT='m110',
  OIL_FORTIFIED='m112',
  IRON_LWEEK='m414',
  supple ='m415', # took any supplements
  FEVER24='m417',
  HEMOGLOBIN_LAB='m432',
  Malaria_test_result='m431',
  WEIGHT='m435',
  HEIGHT='m436',
  is_pregnant='preg',
  agp_HIGH='agp_c1',
  crp_HIGH='crp_c1',
  ANY_INFLAMMATION='agp_crp_c1',
  ANEMIA_ADJUSTED='anemia',
  SFER_microg_L_ADJUSTED_INTERNAL='sf_reg',
  LOW_SFER_microg_L_UNADJUSTED='sf_c2',
  LOW_SFER_microg_L_ADJUSTED='sf_c1',
  VITA_CONTENT_OIL='oil_vita',
  VITA_CONTENT_SUGAR='sugar_vita',
  IODINE_SALT='salt',
  was_fasting='fast',
  ps_folate='fol',
  rbc_folate='rbcf',
  iodine='iod',
  selenium='sel',
  vitamin_b12_gr='vitb12', 
  time_of_day_sampled2= "m430h", 
had_fever='m416',
had_diarrhea= 'm419',
had_malaria= 'm420') 

# Selecting variables of interest:

Malawi_WRA <- Malawi_WRA %>% select(region, sex , crp,  agp, urbanity,survey_cluster1,
household_id1, survey_weight, LINENUMBER, AGE_IN_YEARS, supple, # took any supplements
Malaria_test_result, WEIGHT, HEIGHT, is_pregnant, selenium) 

```


There are `r sum(!is.na(Malawi_WRA$selenium))` entries for plasma Se in the dataset.

Main cleaning assumptions/ decisions:

- **Age**: Included women with age outside the standard range of WRA (15-49yo). 
- **Gender**: One "male" was recoded as "female"
- **Pregnancy**: Pregnant women and those recorded as `NA` were excluded.
- **Malaria test recode**: 2 (0) == Negative, 1 == Positive, all other `NA`.


#### Age

There are some observations (n = `r sum(Malawi_WRA$AGE_IN_YEARS<15)`) that are below the age range for WRA, and none above the range.

In addition, mean (and median) were higher in the northern region compared with southern and center. Whereas, mean/median age were higher in rural areas. We applied survey weights when calculating them. 


#### Gender & Pregnancy

There is one entry labelled as "male" while the dataset should only contain women, and there were `r sum(Malawi_WRA$is_pregnant==1 | is.na(Malawi_WRA$is_pregnant))` women pregnant or with unkown status, whiche were excluded from the dataset. 

```{r}

# Removing the pregnant women and unknown status
Malawi_WRA  <- subset(Malawi_WRA, is_pregnant==0 | is.na(is_pregnant)) 

```

#### Malaria test 

We needed to recode other values than positive/ negative into `NA`, also we did not follow the recode manual (we are not putting 0 for negative, but keeping 2).

```{r}

# Recoding others values into NA 
Malawi_WRA$Malaria_test_result <- ifelse(Malawi_WRA$Malaria_test_result==4| 
                                        Malawi_WRA$Malaria_test_result==6, 
                                         NA, Malawi_WRA$Malaria_test_result)

Malawi_WRA$Malaria_test_result  <- as.factor(Malawi_WRA$Malaria_test_result)

```


This variable has `r sum(is.na(Malawi_WRA$Malaria_test_result) & !is.na(Malawi_WRA$selenium))` missing values which would reduce the sample size for plasma Se. Hence, we need to evaluate whether the information provided would influence the results to decided whether to add it or not to the model. 

Can we identify if malaria is influencing Se status? - We plotted the results, and they didn't seem to have many differences between positive and negative malaria test, but maybe due to the high number of negative tests (see @sec-inflam).

We could run the model with and without the variable and see if there is any difference in the final results. 

#### Supplementation

Similarly, as with the Malaria test, we found high number of missing values which would reduce the sample size for plasma Se by `r sum(is.na(Malawi_WRA$supple) & !is.na(Malawi_WRA$selenium))`, and also only few reported consuming supplement (n=13), thus, we decided to exclude this variable.


#### Height & Weight


```{r}

# Height: Missing values (999 to NA)
Malawi_WRA$HEIGHT[Malawi_WRA$HEIGHT >999] <- NA

# Weight: Missing values (999 to NA)
Malawi_WRA$WEIGHT[Malawi_WRA$WEIGHT >999] <- NA

# Creating BMI variable
Malawi_WRA$BMI<- Malawi_WRA$WEIGHT/(Malawi_WRA$HEIGHT/100)^2

```

-  **Height:** The height distribution was normal with two potential outliers (i.e.,  one woman with a recorded height of 100.2), after recoding the missing values (999) to `NA`, according to the coding standards reported in (ICF. 2018. Demographic and Health Surveys Standard Recode Manual for DHS7. The Demographic
and Health Surveys Program. Rockville, Maryland, U.S.A.: ICF).   

-  **Weight:** Some outliers were found as well, some weights were higher than expected, on the lower bound seems plausible given age and height.

-  **BMI:** We generated BMI variable which also has some outliers due to the high weight of some women (@fig-3). We are not removing those women as the BMI is still plausible.

```{r label = "fig-3", fig.cap = "Distrubution of Body Mass Index (BMI) (kg/m^2) of women in Malawi"}

summaplot(Malawi_WRA$BMI)

```

We can see also that there is a difference in the BMI between rural and urban women. We can see similar trend with the WQ (@sec-wq), which the Northern region in Malawi has higher BMI than the other two regions.  


#### Selenium

Plasma Se median (and mean) values were higher in urban settings than in rural, and the highest in the Southern region. Then, the median dissagregated by region and urbanicity showed that Northern and Southern urban population had the highest plasma Se median concentration, while the Central region were the lowest, particularly for women living in rural settings. 

What it is interesting is that while the women in the Northern-urban areas had higher plasma Se than their southern counterpart, the women in the Southern-rural areas had higher plasma Se those in Northern-rural areas. In addition, women who live in urban setting of the Central region had similar plasma Se median concentration as women in the northern region in rural settings. 

#### Inflammation {#sec-inflam}

According to the NSO (2017) report: Malawi Micronutrient Survey 2015-2016, inflammation was reported around 12% but it was consistent across age groups, residency, and wealth Q. However, malaria prevalence was higher in rural areas and for the different wealth quintiles (WQ), with lowest WQ being 6-times higher than the highest WQ. Since, these two variables may influence Plasma Se status we are including it in the model. 

It could be that this variable is correlated with Malaria, hence by including these two inflammation variables, we may be accounting for the potential influence of the malaria in the plasma Se.



### Socio-economic data (DHSDATA)

We extracted the socio-economic information from the DHS survey, however, we are only interested in the data for women and those with reported Plasma Se values. 


#### Wealth Quintile {#sec-wq}

The NSO (2017) reported significant differences between the number of individuals in each WQ when disaggregated by rural/urban residency. 

When plotted the  WQ per region, we can see that the Central and Southern region have more or less even distribution of the WQ, however the Northern region has predominately highest and higher WQ, accounting for more than 60% of the population in the Northern region (unweighted results). 


#### Education

We could see when those two variables plotted together that most of the people with higher education was in the highest  WQ. Similarly, women in the lowest WQ has no high education.

All education and wealth level have representation in all three regions. Again, when looking at the data we could see that there is a slight "over-reporting" of the wealthiest quintiles. 

Hence, we excluded this variable as it could possibly be confounding effect. 

```{r}

# Wealth quintile 1-5 (lowest to highest)
# Education 0-3 (low to high)

plot(EligibleDHS$wealth_quintile, EligibleDHS$education_level, 
     main="Wealth vs Education",
     xlab="Wealth Q", ylab="Education level", pch=19)

```

Similarly, when we plotted the Education by region, we could see the same trend as in the WQ. 


#### Literacy

This variable was excluded as some of the results were incongruent. For instance,the education and literacy, all the women with secondary education reported literacy (n = `r sum(EligibleDHS$education_level >1 & EligibleDHS$Literacy == 0)`), however, there were women who reported no literacy (0), but secondary education (2).


### Combining DHS datasets: Plasma Se & covariates

According to the study of Phiri et al (2019), the influencing factors for Plasma Se (for WRA) in Malawi were soil type and proximity to Lake Malawi. 

In addition, other authors have reported that socio-economic characteristics may influence Se status. 

Here we are exploring the variability of Plasma Se according to those variables. First within the same dataset.


#### Wealth

Plasma Se values were similar across WQ, with higher variability in the two lowest WQ. 

#### Education level

Plasma Se values were slightly higher in the highest education level, with lower variability, while the opposite was true for the primary education level. 

#### Place of residency

Plasma Se values were similar when aggregated by region, whereas we can start seeing some differences when we are plotting the results by district. Also notice, that the number of observations per district is 1) much smaller and 2) much diverse (from `r min(table(EligibleDHS$sdist))` to `r max(table(EligibleDHS$sdist))`). 

Because we are exploring how much Plasma Se variability is explained by Maize Se concentration, we do not need to apply (or worry about) the weights. Only, we would need to apply them if we want to express some findings at population level, e.g., perc. of population at risk. We are not likely to do that, as 1) it is not the objective of the study/ model, and 2) we are working at certain scales where data will not be representative (e.g., district). 

**Note**: What it would be important to check, it is whether the data is balance (e.g., highly skewed to one side and hence driving the model outcomes). 
We checked wealth by region and it seemed that there is a slight unbalance for the lowest Q in the middle region and towards the highest Q in the Northern region. See above (@sec-wq). Maybe we need to check for confounding effect.  


#### Smoking

Only 3 respondent reported smoking cigarettes, if we want to include this variable, we may need to combine with other smoking variables (including the other variables), although I believe that women who smoke in this context it may be very low.

After merging the datasets, some WQ were missing, hence following recommendations form F.S we used the household WQ for those available to impute it. They were available for 17 of the 39 observation with missing values.

For Plasma Se, we only have values for `r sum(!is.na(EligibleDHS$selenium))`. This will condition the model choice.


### GPS location

It is at the cluster level, hence we need to check the variation within cluster as this will affect our model selection and results. 

We are adding a buffer around each of the GPS locations that account for the 2km and 5km displacement in urban and rural respectively. In that sense, we are assuming that each HHs can be in any place with that buffered area. 

Final cleaned dataset was stored in: `data/inter-output/dhs-gps.rds`



Then, the next step will be to combine with the information on maize Se at different aggregation level. To do so, first we need to "clean the location" as due to the GPS displacement Plasma Se and maize Se values were not co-located. The steps to "co-locate" Plasma Se and Maize Se are performed in the script "00_cleaning-locations.R". 


#### Limitation to DHS-MNS data

  - **Displacement of the GPS coordinates at cluster level**. This led to high chances of missclassification and reducing the accuracy of the results for small area (e.g., EAs). It also led to the reduction of the no. of co-located (maize Se & plasma Se) samples. 


### Boundaries of Malawi {#sec-explo-bound}

In this study, we are trying to identify the optimal level of aggregation of maize grain Se composition data, hence different set of spatial aggregation units (areal units) were generated based on the administrative units of the country (Admin level 3 - EAs, Admin level 2 - districts), and based on household locations. For instance, areas based on the clusters reported in the DHS dataset.

#### Administrative boundaries

We have found several shape files with the boundaries for Malawi (See @sec-bound). After exploring the boundaries datasets (pre-exploration using QGIS version (3.28.3) Firenze). Then, after selecting boundaries 3, we decided to use `ea_bnd` for EAs, and generate from that file (dissolving boundaries) the district boundaries file (`dist_bnd`), to avoid potential effect/ bias in the results due to differences in the boundaries datasets (See below). The dist_bnd file is generated in the script (`00_cleaning-boundaries.R`).


##### Limitation of the admin boundaries data 

The **first issue** that we encountered was that not many of the files have the boundaries at EA level. This is due to the confidentiality and sensitivity of the data, as those are the clusters used for most of the survey and are bases on the census data. 

The **second issue** was that some boundaries are not accurate/ well-defined. For instance, in the dataset 1 (GADM), which is a good source of countries boundaries, lake Malawi was included into the admin boundaries. Hence, when using the boundaries to locate the HHs and/or maize it seemed that part of the household were falling inside Lake Malawi. If we try to remove the lake area it leads to the missing of some observations. 

The **third issue** we found was inconsistency between datasets. Although boundaries 2 and 3 are apparently produced by the NSO the number of TA are different (n=`r length(unique(ea_bnd$TA_CODE))`) and (n=`r length(unique(ta_bnd$ADM3_PCODE))`). The disagreement may be due to the inclusion of sub-divided TA (i.e., TSA) in the dataset 3, or due differences on the date of the datasets. There are are small differences in the codes used between the datasets which may yielded further inconsistencies between the two. Hence, we decided to exclude this boundaries for further analysis and do not aggregate at this level. In addition, after discussing with GO and EJ, they agreed that TA may be inconsistent boundaries and that may have some political/ local sensibilities. 

More info (on practical steps & tested thing on the notebook `Bondaries_notes.md`). 
